---
title: "Covid91 vaccine study Final2024"
author: "Kevin Ortega"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

## Methods

Question: Is the covid91 vaccine effective in males?

$H_0$: The vaccine is not effective in preventing covid91 in males.

$H_A$: The vaccine is effective in preventing covid91 in males.

infected~sex(M)
categorical~categorical

We'll be using Cat~Cat

## Graphical Results

```{r}
Male <- FinalDat[FinalDat$sex=='M',]
```


```{r}
barchartGC(~infected + treatment,data=Male)
barchartGC(~infected + treatment,data=Male, type="percent")

```
From this bar chart it shows that you are less likely to contract covid91 if you did receive the vaccine.

## Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=Male)
rowPerc(table1)
colPerc(table1)
```
The numerical data shows the results in percentages. Within the placebo group 73.80% of individuals got infected with covid91. While in the vaccine group 26.20% contracted covid-91. This indicates the receiving the vaccine as a male significantly reduces the risk of contracted covid-91. 

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

This data and the Fisher exact test shows that the P-Value is 2.2e-16 which is significantly lower than 0.05 which means we reject the null hypothesis. 

## Conclusion

The bar chart shows that you are less likely to contract covid91 if you did receive the vaccine. The numerical data shows the results in percentages. Within the placebo group 73.80% of individuals got infected with covid91. While in the vaccine group 26.20% contracted covid-91. This indicates that receiving the vaccine as a male significantly reduces the risk of contracted covid-91. The inferential data including the Fisher exact test shows that the P-Value is 2.2e-16 which is significantly lower than 0.05 which means we reject the null hypothesis. All in all these examples and simulations show that is as a male you are far less likely to contract covid-91 upon receiving the vaccine.

# Females

## Methods

Question: Is the vaccine effecitve in preventing Covid91 in females?

$H_0$: The vaccine is not effective in preventing covid91 in females.

$H_A$: The vaccine is effective in preventing covid91 in females


infected~sex(F)
categorical~categorical

We'll be using Cat~Cat

## Graphical Results

```{r}
Female <- FinalDat[FinalDat$sex=='F',]
```

```{r}
barchartGC(~infected + treatment,data=Female)
barchartGC(~infected + treatment,data=Female, type="percent")

```
From this chart it shows that females are less likely to contract Covid91 if the recieve the vaccine.

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Female)
rowPerc(table1)
colPerc(table1)
```
The numerical data shows the results in percentages. Within the placebo group 64.93% of individuals got infected with covid91. While in the vaccine group 35.07% contracted covid-91. This indicates the receiving the vaccine as a female significantly reduces the risk of contracted covid-91.

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

This data and the Fisher exact test shows that the P-Value is 1.246e-08 which is significantly lower than 0.05 which means we reject the null hypothesis.

## Conclusion

The bar chart shows that you are less likely to contract Covid91 if you did receive the vaccine. The numerical data shows the results in percentages. Within the placebo group 64.93% of individuals got infected with Covid91. While in the vaccine group 35.07% contracted covid-91. This indicates that receiving the vaccine as a female significantly reduces the risk of contracted Covid-91. The inferential data including the Fisher exact test shows that the P-Value is 1.246e-08 which is significantly lower than 0.05 which means we reject the null hypothesis. In summation these examples and simulations show that is as a female you are far less likely to contract covid-91 if you receive the vaccine.

## LGBTQ

$H_0$: The vaccine is not effective in preventing covid91 in members of the LGBTQ community.

$H_A$: The vaccine is effective in preventing covid91 in members of the LGBTQ community.

## Druggies

$H_0$: The vaccine is not effective in preventing covid91 in drug user.

$H_A$: The vaccine is effective in preventing covid91 in drug users.


# Overall Results and Conclusions